\boldmath Commissioning of the highly granular SiW-ECAL technological prototype

### 2019-04-04 performance-paper-draft/commissioning_arxiv1810.05133/arxiv-v0/draft_arxiv-v0.tex
Up to date to the arxiv submissions

\linenumbers

@@

% e-mail addresses: only for the forresponding author
\emailAdd{irles@lal.in2p3.fr}
%\collaboration[]{on behalf of the CALICE collaboration}

\abstract{
  High precision physics at future colliders as the International Linear Collider (ILC) require unprecedented high precision in the determination
  of the final state of the particles produced in the collisions
  Thia precission will be achieved thanks to the Particle Flow algorithms (PF) which require compact, highly granular and hermetic calorimeters systems.
  The Silicon-Tungsten Electromagnetic Calorimeter (SiW-ECAL) technological prototype
  design and R\&D is oriented at the baseline design of the ECAL of the International Large Detector (ILD) for the ILC.
  In this article we present the commissioning and the performance of the prototype in a beam test carried at DESY in June 2017.
}

@@

Future accelerator based particle physics experiments
require very precise and detailed reconstruction of the final states produced
in the beam collisions. A particular example is the next generation of $e^{+}e^{-}$
linear colliders such the ILC\cite{Behnke:2013xla,Baer:2013cma,Adolphsen:2013jya,Adolphsen:2013kya,Behnke:2013lya}.
This project will provide collisions of polarized beams with center-of-mass energies of 250 GeV - 1 TeV.
These collisions will be studied by two multipurpose detectors:
the International Large Detector (ILD) and the Silicon Detector (SiD)\cite{Behnke:2013lya}.
To meet the precision levels required by the ILC %or CLIC
physics goals,
new techniques relying on single particle separation to make possible the choice of the best information available
in the full detector to measure the energy of the final state objects have been developed.
These techniques are called Particle Flow (PF) techniques \cite{Brient:2002gh,Morgunov:2004ed,Sefkow:2015hna}
and allow to reduce the impact of the poor resolution of the calorimeter systems (compared with trackers) in the overall reconstruction.
The PF algorithms impose some special requirements in the design of the detectors.
For example, it requires highly granular, compact
and hermetic calorimeters.

The CALICE collaboration is driving most of the efforts on R\&D of highly granular calorimeters \cite{Sefkow:2015hna} 
for future linear colliders by investigating and building prototypes for several
calorimeter concepts. One of these calorimeters 
is the silicon-tungsten electromagnetic calorimeter, SiW-ECAL.
The SiW-ECAL is the baseline choice for the ILD electromagnetic calorimeter.
It consists of a detector (in the barrel region) of 24 $X_{0}$ of thickness which corresponds to $\sim 1~\lambda_{I}$ (interaction length).
It has silicon (Si) as active material and tungsten (W) as absorber material.
The combination of Si and W choices  makes possible the design and construction
of a very compact calorimeter with highly granular and compact active layers.
It will be built an alveolar structure of carbon fiber into which modules made of tungsten
plates and the active sensors will be inserted. The very-front-end (VFE) electronics will be
embedded in the detector units. The silicon sensors will be segmented
in squared cells (or channels) of 5x5 mm: a total of $\sim 100$ million readout channels will constitute the ECAL for ILD.
The desired signal dynamic range in each channel goes from 0.5 MIP to 3000 MIPs, where the MIP acronym stands 
for the energy deposited by a minimum-inonizing-particle.
To reduce overall power consumption, the SiW-ECAL will exploit the special bunch structure
foreseen for the ILC: the $e^{+}e^{-}$ bunches trains will arrive in spills
of $\sim$ 1-2 ms width separated by $\sim$ 200 ms. The data acquisition will be gated during these short windows and during the idle time the bias currents of the electronics will be shut down.
This technique is usually denominated power pulsing. In addition to this, to cope with the large amount of channels,
the calorimeters should work in self-trigger mode (each channel featuring an internal trigger decision chain) and zero suppression mode.

@@

The first SiW-ECAL prototype was the so called SiW-ECAL physics prototype.
It was successfully tested at DESY, FNAL and CERN running in front of another prototype from the CALICE
collaboration, the analogue hadronic calorimeter AHCAL, delivering the proof of concept of the technollogy
and the PF calorimetry.
For the physics prototype, the VFE was placed outside the active area with no particular constraints in power consumption.
It consisted of 30 layers of Si as active material alternated with tungsten plates as absorber material.
The active layers were made of a matrix of 3x3 Si wafers of 500 $\mu$m thickness. Each of these wafers was segmented in matrices of
6x6 squared channels of 1x1 $cm^{2}$, allowing for a potential density of 1500 channels/dm$^{3}$ assuming
the ILD baseline design constraints on the material repartition and compactness.
The prototype was divided in 3 modules of 10 layers with different W depth per layer in each of these modules
(0.4, 1.6 and 2.4 $X_{0}$) making a total of 24 $X_{0}$.
That very first prototype offered a signal over noise on the measured charge of 7.5 for MIP like 
particles. More results proving the good performance of the technology and the PF can be found in
references~\cite{Adloff:2011ha,Anduze:2008hq,Adloff:2008aa,Adloff:2010xj,CALICE:2011aa,Bilki:2014uep}.

The current prototype is called the SiW-ECAL technological prototype. It addresses the main technological challenges: compactness,
power consumption reduction through power pulsing and VFE inside the detector close to real ILD conditions.
It will also provide data to deeply study the PF and provide input to tune simulation programs as for example
GEANT4\cite{Agostinelli:2002hh,Allison:2006ve,Allison:2016lfl} which is widely used
in particle physics to simulate the passage of particles through matter. In this section we described in detail
the main features and characteristics of the technological prototype.

@@

The sensors consist of high resistivity (bigger than 5000 $\Omega\cdot$cm)
%floating zone
silicon wafers with a thickness of 320$\mu$m.
The size of the wafers is $9\times9$ cm$^{2}$ and each of them is subdivided in an array of 256 PIN diodes of $5\times5$ mm$^{2}$.
A MIP traversing the PIN parallel to its normal will create $\sim$ 80 $h^{+}e^{-}$ pairs per $\mu$m which corresponds to 4.1 fC
for particles incident perpendicularly to its surface.
The original design of the silicon wafers included an edge termination made of floating guard-rings.
It was observed in beam tests \cite{Cornat:2015eoa,Cornat:2009zz} that the capacitive coupling between such floating guard-rings 
and the channels at the edge created not negligible rates off fake
events in tests with high energy beams (pions and electrons with energies larger than 20-40 GeV)
An R\&D program together with Hamamatsu Photonics (HPK Japan) was conducted to study the guard-rings design 
as well as the internal crosstalk. It was concluded that using wafers without guard rings and with a width of the peripheral areas lower than 
500$\mu$m thanks to the use of stealth dicing technique, the amount of these squared events 
can be reduced to be at negligible level.
For the setup described this article we used different solutions for the edge terminations.
For all of them, the expected levels of fake events are negligible due to
the low energy of the interactions studied in this paper:
MIP like particles or electromagnetic showers created by particles of few GeV
from the DESY beam. Therefore, they are not further discussed here.

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=4in]{../figs/skiroc2_block.eps}
\caption{The schematics of the analog part of SKIROC2. \todo{High-stack picture (right bottom corner)}}
\label{SKIROC2}
\end{figure}

The SKIROC\cite{Callier:2011zz} (Silicon pin Kalorimeter Integrated ReadOut Chip) is a
very front end ASIC (application-specific integrated circuits)
designed for the readout of silicon PIN diodes.
In its version SKIROC2 it consists of 64 channels in AMS 0.35 $\mu$m SiGe technology.
A schematic view of the analog part of the SKIROC2 is shown in Figure \ref{SKIROC2}.
Each channel comprises a low noise charge preamplifier of variable gain followed by two branches:
a fast shaper for the trigger decision and a set of dual gain slow shapers for charge measurement. 
The gain can be controlled by modifying the feedback capacitance during the configuration of the detector.
With the lowest gain, 6pF, the ASIC will handle a linear dynamic range from 0.1 to up to 1500 MIPs. 
Finally, a Wilkinson type analogue to digital converter fabricates the digitized charge deposition that can be readout. 
Once one channel is triggered, the ASIC reads out all 64 channels adding a bit of information to tag them as
triggered or not triggered and the information is stored in a 15 cell deep physical switched capacitor array (SCA).

The SKIROC ASICs can be power-pulsed by taking advantage of the ILC spill structure: 
the bias currents of the ASIC can be shut down during the idle time between bunch trains.
With this method, the ASIC is able to reduce its power consumption down to 25 $\mu$W per channel,
meeting the ILC requirements. 
The power pulsing feature is used for all the results discussed in this paper
and for first time in long periods of data taking in beam test.

@@

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
    \includegraphics[width=4in]{../figs/ASU_geometry1.eps}  \\
    \includegraphics[width=4in]{../figs/ASU_geometry2.eps}  \\
  \end{tabular}
  \caption{Repartition of the ASIC (up) and channels (down) in one ASU. In this perspective, the PCB is in the top
    and the sensors are in glued in the back.
    The channels are separated (in x and y) by 5.5 mm.
    The empty cross in the middle of the ASU corresponds to the 1 mm separation between the sensors.
    The areas covered by the different ASICs and channels
    are labeled with numbers following design and DAQ criteria: from 0-16 in the case of the ASICs and from 0-63 in the case of the channels.
  }
\label{ASU}
\end{figure}

The entity
of sensors, thin PCB (printed circuit boards) and ASICs is called Active Signal Units or ASU.
An individual ASU has a lateral dimension of 18x18 cm$^{2}$.
The ASUs are currently equipped
further with 16 SKIROC2 ASICs for the read out and features 1024 square pads (64 per ASIC) of 5x5 mm.
The channels and ASICs are distributed along the ASU as shown in Figure \ref{ASU}. Each ASU is equipped with 4 silicon wafers as the described in Section \ref{sec:wafers}.
The high voltage is delivered to the wafers using a HV-kapton sheet that covers the full extension of the wafers.

The current version of the PCB is called the FEV11. It has a thickness of 1.6mm which
grows up to 2.7mm when the ASICs in its current packaging (1.1 mm thick LFBGA package) are
bonded in top of it.
With these characteristics, a potential density of
3800 channels/dm$^{3}$ is achievable keeping the space and
interaction length requirements of the the baseline design of the ECAL forthe ILD.
This number should be compared with
the density achieved in beam tests with the physics prototype: 1500 channels/dm$^{3}$.
With the first versions
of the technological prototype we reached similar potential density level as in
the current version but equipping only a
quarter of the ASUs surface \cite{Amjad:2014tha}.

@@

\begin{enumerate}
\item The first component is the so called detector interface (DIF) which is placed at the beginning of each layer holding up to 15 ASUs.
\item All DIFs are connected by single HDMI cables to the concentrator cards as the second component: the Gigabit Concentrator Cards (GDCCs). These cards are used to control up to 7 DIFs. They collect all data from the DIFs and distribute among them the system clock and fast commands.
  \item The most downstream component, is the clock and control card (CCC) which
provides a clock, the control fan-out of up to 8 GDCCs and accepts and distributes external signals (i.e. signals
generated external pulse generator to simulate the ILC spill conditions).
\end{enumerate}

@@

\subsection{Fully equipped readout modules: the SLABs}
\label{sec:setup}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=6in]{../figs/short_slab_bga.jpg} 
  \caption{Open single SLAB with FEV11 ASU, 16 SKIROC 2 and the interface card visibles. }
\label{ASU2}
\end{figure}

A fully equipped readout module is shown in Fig. \ref{ASU2}. These modules are called
SLABs and consist of a chain of one or several fully equipped ASUs
connected to a data acquisition system (DAQ) through an adapter board, called SMBv4.
The SMBv4 also serves as to hold other
services as power connectors or the super capacitances used for the power pulsing. 
These capacitances of 400mF with 16 m$\Omega$ of equivalent serial resistance
are extra dimensioned to provide enough local storage 
of power to assure stable low voltage levels during the power pulsing. %In this configuration
%the high currents are still crossing the interconnections/
The readout modules are embedded on a "U" shaped carbon structure to protect the wafers.
The full system is then covered by two aluminum plates
to provide electromagnetic shielding and mechanical stability.

For the production of the small sample of SLABs studied in this document,
a scalable working procedure has been established among several groups \cite{Boudry:2318814}
profiting from the funding of projects like AIDA2020 or the HIGHTEC emblematic project
of the P2IO. A schematic view of this assembly procedure chain can be seen in
Figure \ref{assembly}. For more details we refer to Ref.\cite{Boudry:2318814}.
This process is to be extrapolated to a full assembly procedure for
e.g. the ILD detector.

\begin{figure}[!t]
\centering
\begin{tabular}{l}
\includegraphics[width=4.0in]{../figs/assembly.png} 
\end{tabular}
\caption{Process flow for the assembly of the SiW-ECAL SLABs.}
\label{assembly}
\end{figure}

\subsection{The prototype setup}
\label{sec:setup2}

A picture showing the SiW-ECAL technological prototype setup can be seen in Figure \ref{proto}.
The current prototype consists on 7 layers of SLABs housed in a PVC and aluminum structure that can host up to 10 layers in slots separated by 15mm each.
The first six layers were placed in the first six slots and the last one was in the last slot,
with respect to the beam pipe. In the following sections, we will refer to layers number 1 to 7, where
the 1 is the closest to the beam pipe and 7 is the farthest.
This setup is used for commissioning (Section \ref{sec:commissioning}) and for the beam test
(Section \ref{sec:beamtest}). In both cases, the detector was
running in power pulsing mode.

\begin{figure}[!ht]
\centering
\begin{tabular}{l}
\includegraphics[width=4.0in]{../figs/proto.png} 
\end{tabular}
\caption{Prototype with 7 layers inside the aluminum stack.}
\label{proto}
\end{figure}

@@

During the commissioning, we observed the repetition of coherent noise events affecting to several
SLABs at the end of acquisitions with long gating time. The situation
could be remedied by improving the isolation of the individual SLABs and by reducing the
data taking to short gating times.
In any case, all runs dedicated to the commissioning are usually characterized by their short gating windows for the acquisition (1-2ms)
at low repetition frequencies (1-5 Hz)
to minimize the chances of having real events due to cosmic
rays during the data taking.

@@

The list of the noisy channels was
obtained by means of dedicated data taking runs. In these runs
we scan relatively high trigger thresholds (between 1-3 MIPs) and progressively mask channels that exhibit counts. 
In each step, the decision of tagging a channel as noisy was taking following the next rules:

\begin{itemize}
\item if the channel was triggered at rates larger than 0.5-1\% of the total number of triggers per ASIC it was added to the list;
\item if a channel was tagged as noisy in, at least, three of the SLABs, it was tagged as noisy for all and added to the list of channels being suspect of suffer from routing issues.
\end{itemize}

Following this procedure, we found to different types of noisy channels. One set consists of
channels randomly distributed along the surface of every ASU and the other consists
of channels located in specific areas and systematically noisy in all the ASUs. Preliminary inspections of the PCB layout
hint that the channels in the latter set may be noisy due to
improvable routing of the PCB.
Deeper studies on the PCB routing must be conducted to clarify this.
All the noisy channels have been identified and masked and the power
of their preamplifiers has been disabled. All the results shown in the following
sections are obtained in these conditions.

In addition to the different noisy channel types described above, we also have
masked full sectors of the SLABs if an ASIC was tagged as faulty (at least 70\% of channels 
listed as noisy) or if a Si-wafer was damaged (high leakage currents).
The results of this study is summarized in Figure \ref{noisycells}.

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
  \includegraphics[width=4in]{../figs/commissioning/masked_layer.eps} \\
  \includegraphics[width=4in]{../figs/commissioning/masked_chip.eps}
  \end{tabular}
\caption{Fraction of channels that are tagged as noisy in all slabs. 
Top: different type of noisy channels per slab. 
Bottom: break down of the total number of noisy channels per ASIC. 
The ASICs 4-7 (wafer issue) and 10 from layer 1 and the ASIC 4 from layer 7 are not included
in the second plot since they are fully masked.}
\label{noisycells}
\end{figure}

@@

After the noisy channels have been masked, dedicated trigger threshold
scan runs are taken, and the results are shown in the threshold scan curves where the
x-axis represents the threshold value and the y-axis
the number of recorded signals normalized to 1. The threshold values are given in internal DAC units
which are translated to meaningful physical quantities in Section \ref{sec:comm_trigger_sn}.
In the absence of external signals (cosmic rays, injected signals, etc) 
the falling edge position in the threshold scan curves
is due to the electronic noise
at the output of the fast shaper (the trigger decision branch on the SKIROC)
and it depends on the slow clock frequency.
These threshold scan curves are approximated by a complementary error function:

\begin{equation}
\frac{2p_{0}}{\sqrt(\pi)} \int_{\frac{DAC-p_{1}}{p_{2}}}^{\infty} e^{-t^{2}} dt,
\label{eq_S-curve}
\end{equation}
where $p_{0}$ is 1/2 of the normalization, $p_{1}$ is the value in which the noise levels are 
the 50\% of its maximum and $p_{2}$ give us the width of the threshold scan curve. 
In Figure \ref{scurve_channels} 
two threshold scans curves are shown together with the fit by
the theoretical function.

>>>>

%\linenumbers

@@

\author{\centering S.\,Bilokin\footnote{ Now at IPHC, Strasbourg.}, J.\,Bonis, P.\,Cornebise, A.\,Gallas, A.\,Irles,  R.\,P\"oschl, F.\,Richard, A.\,Thiebault, D.
\,Zerwas \\ {\it Laboratoire de l'Acc\'elerateur Lin\'eaire, CNRS/IN2P3 et Universit\'e de Paris-Sud XI, Centre Scientifique d'Orsay B\^atiment 200, BP 34, F-91898 Orsay 
CEDEX, France } 
\\ M.\,Anduze, V.\,Balagura, V.\,Boudry, J-C.\,Brient, E.\,Edy, G.\,Fayolle, M.\,Frotin, F.\,Gastaldi, A.\,Lobanov, F.\,Magniette, J.\,Nanni, M.\,Rubio-
Roy \footnote{ Now at Spintec, Grenoble} , K.\,Shpak, H.\,Videau, D.\,Yu\footnote{ Now at IHEP, Beijing.} \\ {\it Laboratoire Leprince-Ringuet (LLR) -- \'{E}cole Polytechnique, 
CNRS/IN2P3, Palaiseau, F-91128 France } 
\\ S.\,Callier, F.\,Dulucq, Ch.\,de la Taille, N.\,Seguin-Moreau \\ {\it Laboratoire OMEGA -- \'{E}cole Polytechnique-CNRS/
IN2P3, Palaiseau, F-91128 France } 
\\ J.E.\,Augustin, R.\,Cornat, J.\,David, P.\,Ghislain, D.\,Lacour, L.\,Lavergne\footnote{ Now at IRAP, Toulouse.}, J.M.\, Parraud \\ {\it Laboratoire de Physique Nucl\'eaire et de Hautes Energies 
(LPNHE), Sorbonne Universit\'e, UPD, CNRS/IN2P3, 4 Place Jussieu, 75005 Paris, France } 
\\ K.\,Kawagoe, Y.\,Miura, I.\,Sekiya, T.\,Suehara, T.\,Yoshioka, \\ {\it Department of Physics and Research Center for Advanced Particle Physics, 
 Kyushu University, 744 Motooka, Nishi-ku, Fukuoka 819-0395, Japan } 
\\ D.\,Jeans \\ {\it Institute of Particle and Nuclear Studies, KEK, 1-1 Oho, Tsukuba, Ibaraki 305-0801, Japan } 
\\ J.\,S.\,Chai \\ {\it Department of Electrical and Computer Engineering, Sungkyunkwan Universtity, 16419, Suwon, Gyeonggi-do, Korea}}

% e-mail addresses: only for the forresponding author
%\collaboration[]{on behalf of the CALICE collaboration}
\emailAdd{irles@lal.in2p3.fr}

\abstract{
In this article we describe the commissioning and a first analysis of the the beam test performance of a small prototype of a highly granular silicon tungsten calorimeter. The prototype features detector elements with a channel number similar to that envisaged for e.g. the ILD Detector of the International Linear Collider (ILC). The analysis demonstrates the capability of the detector to record signals as low as 0.5 MIP. Further, no loss of performance has been observed when operating the detector in a high magnetic field.   
}

@@

Future accelerator based particle physics experiments
require very precise and detailed reconstruction of the final states produced
in the beam collisions. A particular example is the next generation of $e^{+}e^{-}$
linear colliders such the ILC\cite{Behnke:2013xla,Baer:2013cma,Adolphsen:2013jya,Adolphsen:2013kya,Behnke:2013lya}.
This project will provide collisions of polarized beams with center-of-mass energies of 250 GeV to 1 TeV
studied by multipurpose detectors. Two projects are proposed:
the International Large Detector (ILD) and the Silicon Detector (SiD)\cite{Behnke:2013lya}.
To meet the precision required by the ILC %or CLIC
physics goals, event reconstruction is based on the 
Particle Flow (PF) technique \cite{Brient:2002gh,Morgunov:2004ed,Sefkow:2015hna}.
PF implies the reconstruction of every single particle of the final state. 
This in turn requires the construction of calorimeters with unprecedented granularity.

The CALICE collaboration is driving most of the efforts on R\&D of highly granular calorimeters
for future linear colliders by investigating and building prototypes for several
calorimeter concepts. One of these calorimeters 
is the silicon-tungsten electromagnetic calorimeter, SiW-ECAL.
Sensors made of pixellated Silicon diodes (Si) are sandwiched between tungsten (W) absorbers.
The combination of Si and W choices  makes possible the design and construction
of a very compact calorimeter with highly granular and compact active layers.
%30 layers in total
%in the barrel region in modules of 20 cm thick.
%It will be built an alveolar structure of carbon fiber into which modules made of tungsten
%plates and the active sensors will be inserted. 
In the detectors at the ILC, the very-front-end (VFE) electronics of the calorimeters will be
embedded in the detector units. 
%The silicon sensors will be segmented
%in squared cells (or channels) of 5x5 mm: a total of $\sim 100$ million readout channels will constitute the ECAL for ILD
%and a density of channels of 6000 channels/dm$^{3}$ in the barrel.
The desired signal dynamic range in each channel goes from 0.5 MIP to 3000 MIPs, where the MIP acronym stands 
for the energy deposited by a minimum-iononizing-particle.
To reduce overall power consumption, the SiW-ECAL will exploit the special bunch structure
foreseen for the ILC with the $e^{+}e^{-}$ bunches trains grouped in spills
of $\sim$ 1-2 ms width separated by $\sim$ 200 ms. 
The front-end electronics will only be enabled during the spills and the bias currents will be shut down between the spills. 
This technique is usually denominated power pulsing. 
The calorimeters are operated in self-trigger mode (each channel featuring an internal trigger decision chain) and zero suppression mode.

@@

The first SiW-ECAL prototype was the so called SiW-ECAL physics prototype.
It was extensively tested between 2005 and 2011 at DESY, FNAL and CERN ~\cite{Adloff:2011ha,Anduze:2008hq,Adloff:2008aa,Adloff:2010xj,CALICE:2011aa,Bilki:2014uep}. 
%running in front of another prototype from the CALICE
%collaboration, the analogue hadronic calorimeter AHCAL, delivering the proof of concept of the technollogy
%and the PF calorimetry.
For the physics prototype, the VFE was placed outside the active area with no particular constraints in the power consumption.
It consisted of 30 layers of Si as active material alternated with tungsten plates as absorber material.
The active layers were made of a matrix of 3x3 Si wafers of 500 $\mu$m thickness. Each of these wafers was segmented in matrices of
6x6 squared channels of 10x10 mm$^{2}$.
The prototype was divided in 3 modules of 10 layers with different W depth per layer in each of these modules
(0.4, 1.6 and 2.4 $X_{0}$) making a total of 24 $X_{0}$.
That very first prototype offered a signal over noise on the measured charge of 7.5 for MIP like 
particles.

The current prototype is called the SiW-ECAL technological prototype. It addresses the main technological challenges: compactness,
power consumption reduction through power pulsing and VFE inside the detector.
%It will also provide data to deeply study the PF and provide input to tune simulation programs as for example
%GEANT4\cite{Agostinelli:2002hh,Allison:2006ve,Allison:2016lfl} which is widely used
%in particle physics to simulate the passage of particles through matter. 
The study presented here extends considerably the work presented in Ref. \cite{Amjad:2014tha} 
on an earlier version of the technological prototype that comprised 
e.g. four times less channels per layer. 
Further, in this earlier study the detector was still operated in continuous power mode.
In this section we described in detail
the main features and characteristics of the technological prototype.

@@

The sensors consist of high resistivity (bigger than 5000 $\Omega\cdot$cm)
%floating zone
silicon wafers with a thickness of $320\pm15\mu$m.
The size of the wafers is $9\times9$ cm$^{2}$ and each of them is subdivided in an array of 256 PIN diodes of $5\times5$ mm$^{2}$.
A MIP traversing the PIN parallel to its normal will create $\sim$ 80 $h^{+}e^{-}$ pairs per $\mu$m which corresponds to 4.1 fC.
%for particles incident perpendicularly to its surface.
The original design of the silicon wafers included an edge termination made of floating guard-rings.
It was observed in beam tests \cite{Cornat:2015eoa,Cornat:2009zz} that the capacitive coupling between such floating guard-rings 
and the channels at the edge created not negligible rates of parasitic signals.
events in tests with high energy beams (pions and electrons with energies larger than 20-40 GeV)
An R\&D program together with Hamamatsu Photonics (HPK Japan) was conducted to study the guard-ring design 
as well as the internal crosstalk. It was concluded that using wafers without guard-ring and with a width of the peripheral areas lower than 
500$\mu$m thanks to the use of stealth dicing technique, the amount of these squared events 
can be reduced to be at negligible level.
For the setup described in this article we used different solutions for the edge terminations.
For all of them, the expected rate of fake events is however negligible due to
the low energy of the particle beam studied in the present article.
%MIP like particles or electromagnetic showers created by particles of few GeV
%from the DESY beam. Therefore, they are not further discussed here.

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=4in]{skiroc2_block.png}
\caption{The schematics of the analog part of SKIROC2.}
\label{SKIROC2}
\end{figure}

The SKIROC\cite{Callier:2011zz} (Silicon pin Kalorimeter Integrated ReadOut Chip) is a
very front end ASIC (application-specific integrated circuits)
designed for the readout of silicon PIN diodes.
In its version SKIROC2 it consists of 64 channels in AMS 0.35 $\mu$m SiGe technology.
A schematic view of the analog part of the SKIROC2 is shown in Figure \ref{SKIROC2}.
Each channel comprises a low noise charge preamplifier of variable gain followed by two branches:
a fast shaper for the trigger decision and a set of dual gain slow shapers for charge measurement.
The gains can be controlled by modifying the feedback capacitance.
%With the lowest gain, 6pF, the ASIC will handle a linear dynamic range from 0.1 to up to 1500 MIPs. 
When there is a trigger, the charge of the triggered channel is held after a fixed delay 
and stored in one of the memory cells of the 15 cell deep physical switched capacitor array (SCA). 
At the end of the slow clock period in which a trigger occurred
the held of the charge of all the channels not triggered is launched.
Finally, the signals are digitized by a Wilkinson type analogue to digital converter.
The digitized information includes tags to identify the channels with and without trigger.

The SKIROC ASICs can be power-pulsed. 
Under conditions as designed for the ILC the power consumption is expected to be as low as 25muW.
The power pulsing feature is used for all the results discussed in this paper.

@@

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
    \includegraphics[width=4in]{ASU_geometry1-eps-converted-to.pdf}  \\
    \includegraphics[width=4in]{ASU_geometry2-eps-converted-to.pdf}  \\
  \end{tabular}
  \caption{Repartition of the ASIC (up) and channels (down) in one ASU as viewed from the PCB side.
    The channels are separated (in x and y) by 5.5 mm$^{2}$.
    The empty cross in the middle of the ASU corresponds to the 1 mm separation between the sensors.
    The areas covered by the different ASICs and channels
    are labeled with numbers following design and DAQ criteria: from 0-16 in the case of the ASICs and from 0-63 in the case of the channels.
  }
\label{ASU}
\end{figure}

The entity
of sensors, thin PCB (printed circuit boards) and ASICs is called Active Signal Units or ASU.
An individual ASU has a lateral dimension of 18x18 cm$^{2}$.
The ASUs are currently equipped
further with 16 SKIROC2 ASICs for the read out and feature 1024 square pads (64 per ASIC) of 5.5x5.5 mm.
The channels and ASICs are distributed along the ASU as shown in Figure \ref{ASU}. Each ASU is equipped with 4 silicon wafers as the described in Section \ref{sec:wafers}.
The high voltage is delivered to the wafers using a HV-kapton sheet that covers the full extension of the wafers.

The current version of the PCB is called the FEV11. It has a thickness of 1.6 mm which
grows up to 2.7 mm when the ASICs in its current packaging (1.1 mm thick LFBGA package) are
bonded in top of it.
%With these characteristics, a potential density of
%4300 channels/dm$^{3}$ is achievable, to be compared with the channel density of the physics prototype of 
%1500 channels/dm$^{3}$.

@@

\begin{enumerate}
\item The first component is the so called detector interface (DIF) which is placed at the beginning of each layer.
\item Each DIF is connected via a standard HDMI to the  second component: a concentrator card called the Gigabit Concentrator Cards (GDCCs). These cards control up to 7 DIFs. They collect all data from the DIFs and distribute in turn the system clock and fast commands.
  \item The most downstream component, is the clock and control card (CCC) which
provides a clock, the control fan-out of up to 8 GDCCs and accepts and distributes external signals (i.e. signals
generated external pulse generator to simulate the ILC spill conditions).
\end{enumerate}

@@

\subsection{The prototype setup}
\label{sec:setup}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=6in]{short_slab_bga_13.jpg} 
  \caption{Open single readout layer with FEV11 ASU, 16 SKIROC 2 and the interface card visibles. }
\label{ASU2}
\end{figure}

A fully equipped readout module is shown in Fig. \ref{ASU2}. These modules consist in this case of one ASU
connected to a data acquisition system (DAQ) through an adapter board, called SMBv4.
The ASU is the squared PCB on the right side, equipped with the 16 SKIROC2 bump bonded.
It is connected with kapton connectors to the SMBv4 PCB which occupies the rest of the picture.
The SMBv4 also holds other
services e.g. the super capacitance used for the power pulsing. 
This decoupling capacitor (seen in the left bottom part of the photograph)
 of 400mF with 16 m$\Omega$ of equivalent serial resistance
provides enough local storage 
of power to assure stable low voltage supply during the power pulsing. %In this configuration
%the high currents are still crossing the interconnections/
The readout modules are embedded on a "U" shaped carbon structure to protect the wafers.
The full system is then covered by two aluminum plates
to provide electromagnetic shielding and mechanical stability.

For the production of the small sample of readout modules studied in this document,
a scalable working procedure has been established among several groups \cite{Boudry:2318814}
profiting from the funding of projects like AIDA2020 or the HIGHTEC emblematic project
of the P2IO. A schematic view of this assembly procedure chain can be seen in
Figure \ref{assembly}. For more details we refer to Ref.\cite{Boudry:2318814}.
%This process is to be extrapolated to a full assembly procedure for
%e.g. the ILD detector.

\begin{figure}[!t]
\centering
\begin{tabular}{l}
\includegraphics[width=6.0in]{SiW-ECAL-Process_Flow_v2018.pdf}
\end{tabular}
\caption{Process flow for the assembly of the SiW-ECAL readout layers.}
\label{assembly}
\end{figure}

%\subsection{The prototype setup}
%\label{sec:setup2}

A picture showing the SiW-ECAL technological prototype setup can be seen in Figure \ref{proto}.
The current prototype consists of 7 layers of 
readout layers housed in a PVC and aluminum structure that can host up to 10 layers in slots separated by 15 mm each.
The first six layers were placed in the first six slots and the last one was in the last slot.
In the following sections, we will refer to layers number 1 to 7, where
the 1 is the most upstream of the beam.
This setup is used for commissioning (Section \ref{sec:commissioning}) and for the beam test
(Section \ref{sec:beamtest}).

\begin{figure}[!ht]
\centering
\begin{tabular}{l}
\includegraphics[width=4.0in]{proto.png} 
\end{tabular}
\caption{Prototype with 7 layers inside the aluminum stack.}
\label{proto}
\end{figure}

@@

During the commissioning, we observed the repetition of coherent noise events affecting to several
readout layers at the end of acquisitions with long gating time. The situation
could be remedied by improving the isolation of the individual readout layers and by reducing the
data taking to short gating times.
In any case, all runs dedicated to the commissioning are usually characterized by their short gating windows for the acquisition (1-2ms)
at low repetition frequencies (1-5 Hz)
to minimize the chances of having real events due to cosmic
rays during the data taking.

@@

The list of the noisy channels was
obtained by means of dedicated data taking runs. In these runs
we scan relatively high trigger thresholds from larger to lower values ({\it i.e.} between 1-3 MIPs) and progressively mask channels that exhibit counts. 
In each step, the decision of tagging a channel as noisy was taking following the rules described next:

\begin{itemize}
\item if the channel was triggered at rates larger than 0.5-1\% of the total number of triggers per ASIC it was added to the list;
\item if a channel was tagged as noisy in, at least, three of the readout layers, it was tagged as noisy for all and added to the list of channels being suspect of suffer from routing issues.
\end{itemize}

Following this procedure, we found two different types of noisy channels. One set consists of
channels randomly distributed along the surface of every ASU and the other consists
of channels located in specific areas and systematically noisy in all the ASUs. Preliminary inspections of the PCB layout
hint that the channels in the latter set may be noisy due to
improvable routing of the PCB.
Deeper studies on the PCB routing must be conducted to clarify this.
All the noisy channels have been identified and masked and the power
of their preamplifiers has been disabled. All the results shown in the following
sections are obtained in these conditions.

In addition to the different noisy channel types described above, we also have
masked full sectors of the readout layers if an ASIC was tagged as faulty (at least 70\% of channels 
listed as noisy) or if a Si-wafer was damaged (high leakage currents).
The results of this study are summarized in Figure \ref{noisycells}.

\begin{figure}[!t]
  \centering
  \includegraphics[width=4in]{masked_layer-eps-converted-to.pdf} \\
\caption{Fraction of channels that are tagged as noisy in all readout layers.}
\label{noisycells}
\end{figure}

@@

After the noisy channels have been masked, dedicated trigger threshold
scans are performed. The results are presented with the so-called the threshold scan curves where the
x-axis represents the threshold value and the y-axis
the number of recorded signals normalized to 1. The threshold values are given in internal DAC units
which are translated to meaningful physical quantities in Section \ref{sec:comm_trigger_sn}.
In the absence of external signals (cosmic rays, injected signals, etc) 
the falling edge position in the threshold scan curves
is due to the electronic noise
at the output of the fast shaper (the trigger decision branch on the SKIROC)
and it depends on the slow clock frequency.
These threshold scan curves are approximated by a complementary error function:

\begin{equation}
\frac{2p_{0}}{\sqrt(\pi)} \int_{\frac{DAC-p_{1}}{p_{2}}}^{\infty} e^{-t^{2}} dt,
\label{eq_S-curve}
\end{equation}
where $p_{0}$ is 1/2 of the normalization, $p_{1}$ is the value in which the noise levels are 
the 50\% of its maximum and $p_{2}$ give us the width of the error function. 
In Figure \ref{scurve_channels} 
two threshold scans curves are shown together with the fit by
the theoretical function.

\begin{figure}[!ht]
  \centering
  \begin{tabular}{ll}
  \includegraphics[width=2.8in]{scurve_chn18_asic7_layer2-eps-converted-to.pdf} & \includegraphics[width=2.8in]{scurve_chn2_asic0_layer2-eps-converted-to.pdf}
  \end{tabular}
\caption{Two threshold scan curves.}
\label{scurve_channels}
\end{figure}
